\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-06-26 forwards-and-backwards.tex
rv not fitness

% \comment{I'm a bit surprised by the factor of 2 (that manifests as $\varphi^2$),
%     but I've checked where it comes from pretty carefully, and don't see anything wrong.}
Long-term fitness of an individual
is proportional to the fraction of lineages from the distant future that pass through the individual,
and hence the total long-term fitness at a location is proportional to
the stationary distribution of $Y$ there, if it exists.
Therefore,
if $\pi$ is integrable then the per-capita long-term fitness of an individual at $x$
is proportional to $\pi(x) / \varphi(x)$.

@@

Can these two situations be distinguished from summaries of lineage movement?
The first has lineage generator
\[
    f \mapsto \Lgen f = r \gamma \left( \Delta f + 2 \grad \log(\gamma w) \cdot \grad f \right),
\]
while the second has lineage generator $f \mapsto \lambda(x) \Lgen f(x)$.
In other words,
although the stationary profile of the population is unchanged when we scale
local establishment and death by $\lambda$,
the motion of lineages is sped up locally by $\lambda$.
This corresponds to making areas with $\lambda > 1$ more ``sink-like'' 
and $\lambda < 1$ more ``source-like'':
if $\lambda(x) > 1$, then at $x$ both the death rate and probability of 
establishment of new individuals are higher.
As a result, lineages in the second model spend more time in areas with $\lambda < 1$,
i.e., those areas have higher reproductive value, %long-term fitness
something that is, in principle, discernible from genetic data
(because, for instance, making reproductive value less evenly distributed
reduces long-term genetic diversity).

>>>>

% \comment{I'm a bit surprised by the factor of 2 (that manifests as $\varphi^2$),
%     but I've checked where it comes from pretty carefully, and don't see anything wrong.}
The long-term reproductive value of an individual
is proportional to the fraction of lineages from the distant future that pass through the individual,
and hence the total long-term reproductive value at a location is proportional to
the stationary distribution of $Y$ there, if it exists.
Therefore,
if $\pi$ is integrable then the per-capita long-term reproductive value of an individual at $x$
is proportional to $\pi(x) / \varphi(x)$.

@@

Can these two situations be distinguished from summaries of lineage movement?
The first has lineage generator
\[
    f \mapsto \Lgen f = r \gamma \left( \Delta f + 2 \grad \log(\gamma w) \cdot \grad f \right),
\]
while the second has lineage generator $f \mapsto \lambda(x) \Lgen f(x)$.
In other words,
although the stationary profile of the population is unchanged when we scale
local establishment and death by $\lambda$,
the motion of lineages is sped up locally by $\lambda$.
This corresponds to making areas with $\lambda > 1$ more ``sink-like'' 
and $\lambda < 1$ more ``source-like'':
if $\lambda(x) > 1$, then at $x$ both the death rate and probability of 
establishment of new individuals are higher.
As a result, lineages in the second model spend more time in areas with $\lambda < 1$,
i.e., those areas have higher reproductive value,
something that is, in principle, discernible from genetic data
(because, for instance, making reproductive value less evenly distributed
reduces long-term genetic diversity).
